How I use R Markdown to write research papers

### 2020-11-10 text/20/markdown-papers.Rmd
Added experience-not-quantified

---
title: "How I use R Markdown to write research papers"
source: "New"
categories:
- Blog post
tags:
- Writing research papers
output: html_document
---

@@

```{r, echo=FALSE}
suppressMessages(suppressWarnings(library(tidyverse)))
x=seq(-1, 1, length=10000)
y <- sin(1/x)
data.frame(x, y) %>%
  ggplot(aes(x, y)) + geom_line()
```

>>>>

---
title: "How I use R Markdown to write research papers"
source: "New"
categories:
- Blog post
tags:
- Writing research papers
output: html_document
bibliography: rmarkdown-papers.bib
csl: biomed-central.csl
nocite: |
  @ziemann-2016, @zeeberg-2004
---

@@

```{r, echo=FALSE, fig.width=6, fig.height=3}
suppressMessages(suppressWarnings(library(broom)))
suppressMessages(suppressWarnings(library(knitr)))
suppressMessages(suppressWarnings(library(tidyverse)))
x=seq(-1, 1, length=10000)
y <- x*sin(1/x)
data.frame(x, y) %>%
  ggplot(aes(x, y)) + geom_line()
```

With the ability to plot polygons, line segments, and arrows, you can create informational diagrams like this.

```{r, echo=FALSE}
sz <- 1.2
xc <- 20
yc <- 80
xbox <- 10*c(-1, -1,  1,  1, -1) + xc
ybox <- 10*c(-1,  1,  1, -1, -1) + yc
data.frame(x=xbox, y=ybox) %>%
  ggplot(aes(x, y)) +
  expand_limits(x=c(0, 100), y=c(0, 100)) +
  theme_void() +
  geom_path(size=sz) + 
  geom_text(x=xc, y=yc, label="S") +
  geom_path(x=xbox+30, y=ybox-30, size=sz) +
  geom_text(x=xc+30, y=yc-30, label="I") +
  geom_path(x=xbox+60, y=ybox-60, size=sz) +
  geom_text(x=xc+60, y=yc-60, label="R") +
  geom_curve(
    x=xc+40, y=yc-35, xend=xc+55, yend=30, curvature=-0.3, size=sz, 
    arrow = arrow(type = "closed", length=unit(0.15, "inches"))) -> sir_plot
sir_plot +
  geom_curve(
    x=xc+10, y=yc- 5, xend=xc+25, yend=yc-20, curvature=-0.3, size=sz, 
    arrow = arrow(type = "closed", length=unit(0.15, "inches")))
```

The code to produce this graph is a bit tedious to create, but since the code that produces this graph is part of the R Markdown file, you can edit it as you are editting other parts of your paper. If you want to relabel the middle box as an "E", you do not need to exit and run a different program, and then cut-and-paste back into your documents.

**Formulas**

You can get really nice looking mathematical formulas inserted into your paper using a really simple syntax. For example, the code

V = \\frac{4}{3} \\pi r^3

becomes

$V = \frac{4}{3} \pi r^3$

when you sandwich it between two dollar signs.

**Tables**

I used to spend a lot of time retyping numbers from what the statistical software said into a table that was organized to make it more readable. That is a tedious and error prone practice, and if your data set changes, even just a little bit, you have to re-type everything all over again. You can produce nice looking tables in R Markdown. It takes about as much work as just retyping everything, but once the table looks nice, you can use the code over again for other projects. I am still learning how to do tables well, but here is some output that I generated to show the potential value.

```{r}
lm(breaks ~ wool*tension, data = warpbreaks) %>%
  tidy %>%
  rename(t.test=statistic) %>% 
  kable(
    digits=c(1, 1, 1, 2, 4),
    caption="Analysis of variance table")
```

I can change how the values in this table are rounded or the names in the header row with just one line of code. Other enhancements may take a bit more work.

**Bibliographies**

If you have your references stored in BibTeX format, you can place them at the end of your document.
